---
title: "Analysis on Credit card customers"  
output: html_document
---
 *Anumeha* *Mishra - CS 544 Term Project*
 *December 7, 2020*
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 
This dataset has been taken from kaggle and the link is as below: https://www.kaggle.com/sakshigoyal7/credit-card-customers

This dataset contains personal information of the customers such as their age, marital status, income,etc.

This dataset contains 10,128 rows and 20 columns.

The columns attributes are :

CLIENTNUM, Attrition flag,Customer_Age , Gender , Dependent_count , Education_Level, Marital_Status, Income_Category, Card_Category, Months_on_book, Total_Relationship_Count, Months_Inactive, Contacts_Count,Credit_Limit,Total_Revolving_Bal, Total_Amt_Chng_Q4_Q1,Total_Trans_Amt,Total_Trans_Ct,Total_Ct_Chng_Q4_Q1,Avg_Utilization_Ratio

## Objective

The objective of this project is to gain an insight on customers banking and show visualizations on different trends based on their age, gender and marital status.


## Preparing the data

You should consider writing a function whenever you've copied and pasted a block of code more than twice (i.e. you now have three copies of the same code). For example, take a look at this code. What does it do?

```{r}
data <- as.data.frame(read.csv("/Users/anumehamishra/Desktop/BankChurners.csv", header=T, sep=",", na.strings = c("","NA")))
head(data)
```


#### Analyzing the data
##### PART 1
###### Categorical 1- Finding the number of males and females customers
```{r}
table(data$Gender)
barplot(table(data$Gender), col = "cyan", ylim = c(0,6000), las = 2, xlab = "Gender",ylab = "Number of customers")
```


###### Finding the martial status distributions of the customers


```{r}
table(data$Marital_Status)
barplot(table(data$Marital_Status), col = "cyan", ylim = c(0,5000), las = 2, xlab = "Marital_Status",ylab = "Number of customers")
```


###### Numerical 1- Finding the number of males and females customers who have dependents as = 5 


```{r}
Dependents <- data[data$Dependent_count == '5', ]
table(Dependents$Gender)
barplot(table(Dependents$Gender), col = "red", ylim = c(0,250), density=c(10,20),las = 2, xlab = "Dependents = 5", ylab = "Number of customers")
```


##### PART 2
###### Categorical 2- Calculating distribution of  education level among customers

```{r}
Education_Level <- table(data$Education_Level)
slice.labels <- names(Education_Level)
slice.percents <- round(Education_Level/sum(Education_Level)*100)
slice.labels <- paste(slice.labels, slice.percents)
slice.labels <- paste(slice.labels, "%", sep = "")
pie(Education_Level, labels = slice.labels, col = hcl(c(0, 60, 120)))
```


##### PART 3
###### Numerical -  numeric data and analyzing distribution on it 


##### a.  Finding the summary of my data on the basis of Age

###### Boxplot showing the ages of customers 

```{r}
fivenum(data$Customer_Age)
boxplot(data$Customer_Age, horizontal = TRUE, xaxt = "n", xlab = "Age of customers", col=hcl(1))
axis(side = 1, at=fivenum(data$Customer_Age), labels = TRUE)
```


##### b. This will show the distribution of status of customers based on thier age.
```{r}
age_status <- fivenum(data$Customer_Age)
age_status
library(ggplot2)
ggplot(data, aes(x=Customer_Age)) +
  geom_histogram(color="black", fill="blue") + facet_grid(~Attrition_Flag)
```



#####  c- The distriution of age varying with 5 dependents
```{r}
Dependents <- data[data$Dependent_count == '5',]
head(Dependents)
A <- Dependents$Customer_Age
B <- Dependents$Customer_Age
boxplot(A,B, xaxt = "n", xlab = "variations of age with 5 dependents", ylab = "Age of Customers", col=c("yellow", "green"))
```



#####  d- Scatter Plot- Showing age of the customers, count of Dependents and the count of transactions
```{r}
data2=data.frame(Age=data$Customer_Age,Dependent=data$Dependent_count,Transactions=data$Total_Trans_Ct)
head(data2,10)
plot(data2 , pch=20 , cex=1.0 , col="lightblue")
```


#####  e- Plotted a ggplot for age vs Dependent_count, and we have the points colored based on Education_Level
```{r}
ggplot(data = data) +
  geom_point(mapping = aes(x = Customer_Age, y = Dependent_count, colour = Education_Level ))
library(plotly)
```


#####  f - Mixed plot of age, months on book in the bank  and contacts made by each customer
```{r}
a <- plot_ly(data, x = data$Customer_Age, type="box", name = 'age')
b <-add_trace(a, x = data$Months_on_book, type="box", name = 'Months')
c <-add_trace(b , x = data$Contacts_Count,type = "box" , name ="Contacts_made" )
c
```


#####  g - Analysis of distribution on Customers Age
```{r}
values <- data$Customer_Age
tab <- table(values)
dframe <- as.data.frame(tab)
dframe
x <- as.numeric(as.character(dframe$values))
x
```

The probability distribution is:
```{r}
f <- dframe$Freq / (sum(dframe$Freq))
```


The mean:
```{r}
mu <- sum(x * f)
mu
```


Variance of the distribution is:
```{r}
sigmaSquare <- sum((x - mu)^2 * f)
sigmaSquare
sigma <- sqrt(sigmaSquare)
sigma
```

Showing probability distribution of age:


```{r}
plot(x, f, type = 'h', xlab = "Age", ylab = "PMF", ylim = c(0, 0.06), main
     = "Spike plot for Age")
abline(h = 0 )
```


Showing cumulative  distribution of age:


```{r}
cdf <- c(0, cumsum(f))
cdfplot <- stepfun(x, cdf)
plot(cdfplot, verticals=FALSE, pch=16, main="CDF Plot for Age", xlab = "Age", ylab = "C
DF")
```

#####  h - Analysis on the distribution of Customer's income
```{r}
values2 <- data$Dependent_count
tab <- table(values2)
dframe <- as.data.frame(tab)
dframe
x <- as.numeric(as.character(dframe$values2))
x
```

The probability distribution is:
```{r}
f <- dframe$Freq / (sum(dframe$Freq))
```

The mean:
```{r}
mu <- sum(x * f)
mu
```

The variance of the distribution is:
```{r}
sigmaSquare <- sum((x - mu)^2 * f)
sigmaSquare
sigma <- sqrt(sigmaSquare)
sigma
```

Showing probability distribution of Income category:


```{r}
plot(x, f, type = 'h', xlab = "Income_Category", ylab = "PMF", ylim = c(0, 0.5), main
     = "Spike plot for Income")
abline(h = 0 )
```

Showing cumulative  distribution of Income category
```{r}
cdf <- c(0, cumsum(f))
cdfplot <- stepfun(x, cdf)
plot(cdfplot, verticals=FALSE, pch=16, main="CDF Plot for Income", xlab = "Age", ylab = "C
DF")
```

##### PART 4- Central Limit Theorem 
#####  Applying Central Limit Theorem  on age 
```{r}
Age <-data$Customer_Age
ctable <- table(Age)
```

The mean is :
```{r}
mu <- mean(Age)
mu

sigma <- sd(Age)
sigma
```

Histogram showing the age of the customers: 


```{r}
dframe <- as.data.frame(ctable)
x <- as.numeric(as.character(data$Customer_Age))
head(x)
hist(x, probability = TRUE, xlim = c(0, 80), xlab = "Age", ylab = "Density", main = "Histogram of age")
```

#####  Showing results with sample size 5
```{r}
samples <- 1000
sample_size <- 5
xbar <- numeric(samples)
for(i in 1:samples){
  xbar[i] = mean(sample(x, size = sample_size, replace = T))
}
hist(xbar, prob = T, xlab = "Age",
     main = "Densities of age with sample size 5", col = "purple")
```


Mean:
```{r}
mean1 <- mean(xbar)
mean1
```


Standard deviation
```{r}
sd1 <- sd(xbar)
sd1
```
#####  Showing results with sample size 20
```{r}
samples <- 1000
sample_size <- 20
xbar <- numeric(samples)
for(i in 1:samples){
  xbar[i] = mean(sample(x, size = sample_size, replace = T))
}
hist(xbar, prob = T, xlab = "Age",
     main = "Densities of age with sample size 20", col = "red")
```


Mean:
```{r}
mean2 <- mean(xbar)
mean2
```
Standard deviation
```{r}
sd2 <- sd(xbar)
sd2
```
#####  Showing results with sample size 50
```{r}
samples <- 1000
sample_size <- 50
xbar <- numeric(samples)
for(i in 1:samples){
  xbar[i] = mean(sample(x, size = sample_size, replace = T))
}
hist(xbar, prob = T, xlab = "Age",
     main = "Densities of age with sample size 50", col = "brown")
```


Mean:
```{r}
mean3 <- mean(xbar)
mean3
```

Standard deviation
```{r}
sd3 <- sd(xbar)
sd3
```


```{r}
cat("1st distribution:\nMean =",mean1,"\nSD =",sd1)
cat("2nd distribution:\nMean =",mean2,"\nSD =",sd2)
cat("3rd distribution:\nMean =",mean3,"\nSD =",sd3)
```


##### INFERENCE: As we change our sample space and keep increasing it, we notice that the mean almost remains the same whereas the Standard deviation decreases hence proving the property of Central Limit Theorem



##### PART 5- Sampling
```{r}
library(sampling)
table(data$Customer_Age)
hist(data$Customer_Age,col="yellow")
```

##### Using sampling methods for sample size of 200

###### Without replacement 
```{r}
sample.size <- 200
s <- srswor(sample.size,nrow(data))
sample.1 <- data[s != 0, ]
mean_srswor <- mean(sample.1$Customer_Age)
set.seed(153)
```


###### With replacement 
```{r}
s <- srswr(sample.size, nrow(data))
sample.2 <- data[s != 0, ]
mean_srswr <- mean(sample.2$Customer_Age)
set.seed(123)
```

###### Systematic Sampling
```{r}
N <- nrow(data)
n <- 200
```

###### items in each group
```{r}
k <- ceiling(N / n)
k
```


###### random item from first group 
```{r}
r <- sample(k, 1)
r
```

###### selecting every kth item 
```{r}
s <- seq(r, by = k, length = n)
head(s)
```


```{r}
sample.3 <- data[s, ]
table(sample.3$Customer_Age)
mean_systematic <- mean(sample.3$Customer_Age)
mean_systematic
```

###### Systematic sampling with unequal probabilities

```{r}
pik <- inclusionprobabilities(data$Customer_Age, sample.size)
s <- UPsystematic(pik)
sample.4 <- data[s != 0, ]
table(sample.4$Customer_Age)
```

######  Stratified Sampling


```{r}
data["age_range"] = NA
data$age_range <- cut(data$Customer_Age, breaks = c(0, 25, 50, 75, Inf), labels = c('A', 'B', 'C','D'))
data_age <- data.frame(
  Customer_Age = data$Customer_Age,
  age_range = data$age_range
)
freq <- table(data_age$age_range)
freq
```

```{r}
set.seed(123)
head(data_age)
```


###### Cluster sampling
```{r}
cl <- cluster(data, c("Customer_Age"), size = 4, method = "srswor")
sample.6 <- getdata(data, cl)
table(sample.6$Customer_Age)
mean_cluster <- mean(sample.6$Customer_Age)
mean_cluster
```


##### Mean without replacement
```{r}
mean_srswor
```


##### Mean with replacement
```{r}
mean_srswr
```

##### Systematic Mean
```{r}
mean_systematic
```


##### Using sampling methods for sample size of 700


###### Without replacement 
```{r}
sample.size <- 700
s <- srswor(sample.size,nrow(data))
sample.1 <- data[s != 0, ]
mean_srswor <- mean(sample.1$Customer_Age)
```


###### With replacement
```{r}
set.seed(153)
s <- srswr(sample.size, nrow(data))
sample.2 <- data[s != 0, ]
mean_srswr <- mean(sample.2$Customer_Age)
```


###### Systematic Sampling
```{r}
N <- nrow(data)
n <- 700
```

###### items in each group
```{r}
k <- ceiling(N / n)
k
```


###### random items in first group 
```{r}
r <- sample(k, 1)
r
```


###### selecting every kth item
```{r}
s <- seq(r, by = k, length = n)
head(s)
```

```{r}
sample.3 <- data[s, ]
table(sample.3$Customer_Age)
mean_systematic <- mean(sample.3$Customer_Age)
```


###### Systematic Sampling with unequal probabilities
```{r}
pik <- inclusionprobabilities(data$Customer_Age, sample.size)
s <- UPsystematic(pik)
sample.4 <- data[s != 0, ]
table(sample.4$Customer_Age)
```


######  Stratified sampling
```{r}
freq <- table(data_age$age_range)
freq
set.seed(123)
head(data_age)
```


###### Cluster mean
```{r}
mean_cluster
```

###### Systematic mean
```{r}
mean_systematic
```

###### Mean With replacement
```{r}
mean_srswr
```

###### Mean Without replacement
```{r}
mean_srswor
```